### 2013-05-23 milestone.tex
Implemented cross validation for KNN. Began milestone writeup.

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

\title{Exploration of Digit Recognition Algorithms}

\author{
Harnoor Singh \\
Department of Computer Science \& Engineering\\
University of Washington \\
\texttt{hsingh@cs.washington.edu} \\
\And
Brian Walker \\
Department of Computer Science \& Engineering \\
University of Washington \\
\texttt{bdwalker@cs.washington.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

\section{Objective}

As described in our project proposal, we are using the Kaggle digit dataset to classify written numbers.

Our goal is to implement digit recognition using K Nearest Neighbors (KNN) and a Support Vector Machine (SVM) and compare the two approaches. A stretch objective is to enable data that isn't a part of the Kaggle dataset to work with our algorithms. This would require some form of normalization in the input image for consistency.

\section{Progress}

\subsection{KNN}

We have made significant progress in both the KNN algorithm and the SVM algorithm. Our KNN algorithm is completely implemented as is cross validation for it.

To compute the K nearest neighbors, we take the euclidean difference between the feature we are classifying, $X_1$, and every feature in our training set.

\begin{center}
$\sum\limits_{X_2 \exists N} {(X1 - X2)^2}$
\end{center}

Once we have calculated the K closest features, a simple voting method is used to classify $X_1$.

In order to select a value of K we do cross validation on a subset of the training set. We begin by passing a list of potential K values we are considering. For each item, $K_i$ in this list, we hold out $K_i$ samples from our training set and construct our classifier. We then attempt to classify the $K_i$ samples and record the error rate.

The K value that led to the lowest error rate is used to build our final classifier.

\subsection{SVM}

TODO

\begingroup
\renewcommand{\section}[2]{}%
%\renewcommand{\chapter}[2]{}% for other classes
\bibliographystyle{unsrt}
\bibliography{references}	
\endgroup

\end{document}

### 2013-05-23 Digit Recognition/KNN.py

@@ -42,6 +42,8 @@ class KNN:
         return classification
 
     # Store the model in a dictionary
+    # train_data is a matrix of where the first column holds the 
+    # Y values and the other columns are features.
     def __init__(self, train_data, k=1, report=True):
         self.k = k
         self.train_data = train_data

### 2013-05-23 Digit Recognition/Main.py

@@ -5,6 +5,7 @@ import Test
 import Train
 import sys
 import Kernels
+import KNNCrossValidation as cv
 
 output = None
 train_input = None
@@ -12,6 +13,7 @@ test_input = None
 train_data = None
 test_data = None
 test_y = None
+version = None
 
 
 def loadTrainData():
@@ -20,20 +22,22 @@ def loadTrainData():
 
 
 def loadTestData():
-    global test_data, test_input, test_y
+    global test_data, test_input, test_y, version
     test_data = np.loadtxt(test_input, delimiter=",", skiprows=1)
-    test_y = np.loadtxt("./data/knn_benchmark.csv")
+    if version.upper() == "SVM":
+        test_y = np.loadtxt("./data/knn_benchmark.csv")
 
 
 def main():
-    if len(sys.argv) < 4:
-        print "usage: train_file test_file output_directory"
+    if len(sys.argv) < 5 or (sys.argv[4] != "KNN" and sys.argv[4] != "SVM"):
+        print "usage: train_file test_file output_directory version"
         quit()
 
-    global train_input, test_input, output, y, test_data, test_y
+    global train_input, test_input, output, y, test_data, test_y, version
     train_input = sys.argv[1]
     test_input = sys.argv[2]
     output = sys.argv[3]
+    version = sys.argv[4]
 
     loadTrainData()
     loadTestData()
@@ -45,15 +49,24 @@ def main():
     #trainer.trainModel(report=True)
 
 
+    ####################################
+    # SVM DRIVER
     # svm = SVM.Classifier()
     # tester = Test.TestModel()
     # kernel = Kernels.exactPolyKernel(5)
     # model = svm.trainModel(train_data, penalty=1, kernel=kernel, report=True)
     # predictions = tester.predictWithModel(model, test_data)
+    ####################################
 
+
+    ####################################
+    # KNN DRIVER
+    #
+    # Cross validation
+    k = cv.crossValidate([1,2,3,4,5,10,15,20], train_data[0:1000,)
     output = open("./data/output.txt", "w+")
-
-    knn = KNN.KNN(train_data, 1)
+    
+    knn = KNN.KNN(train_data, k)
     classifications = list()
     for i in range(0, len(test_data)):
         x_row = test_data[i, :]
@@ -64,6 +77,11 @@ def main():
     result = sum([np.sign(x) for x in np.subtract(test_y, classifications)])
     print result
     output.close()
[…]

### 2013-05-24 milestone.tex
Updating milestone writeup

@@
 
-\title{Exploration of Digit Recognition Algorithms}
+ \title{Digit Recognition Using KNN and SVM}
 
@@
 
-The K value that led to the lowest error rate is used to build our final classifier.
+The K value that led to the lowest error rate is used to build our final classifier. The result of our cross validation can be seen in Table 1 below.
+
+\begin{table}[ht]
+\caption{Cross Validation Selection of K} % title of Table
+\centering  % used for centering table
+\begin{tabular}{c c} 
+\hline\hline
+K & Error \\ [1.0ex]
+\hline 
+1 & 135 \\
+2 & 164 \\
+3 & 137 \\
+4 & 148 \\
+5 & 144 \\
+10 & 163 \\
+15 & 176 \\
+20 & 191 \\ [1ex]
+\hline
+\end{tabular}
+\label{table:nonlin}
+\end{table}
 
@@
 
-TODO
+The SVM implementations we have worked with previously have involved only two classes. Because we have to classify the digits 0-9, we need to create an SVM that can classify from a broader range.
+
+We are in the process of implementing a one-vs-one classifier.
+We train $n choose 2$ SVMs which compare two numbers.
+For example, one SVM would classify a digit as either a 1 or a 2.
+Another would classify a digit as a 1 or a 3, and so on.
+We get a prediction from all of these classifiers and take a vote to see which digit was most likely based on our SVMs.
+
+In our training step, every time we make a mistake we log it. Future predictions are made based on past mistakes.
+
+\begin{center}
+$Classification = sign(C * \sum\limits_{m \exists Mistakes} {(Y_m * k(X_m, X))})$
+\end{center}
+
+The biggest challenge thus far has been deciding how to regularize our SVM without relying on the weights. Because we use the kernel trick, the weights are never explicitly stored.
+
+\section{Future Work}
+
+Our primary focus moving forward is to get our SVM implementation working.
+Most of the structure is in place, however we need to decide on our method of regularization.
+Once we have the regularization done we can analyze the effectiveness of our SVM.
+
+For KNN it might be beneficial to attemp to increase performance.
+Classifying each sample takes about 1 second, and with the testing dataset holding about 50,000 samples it takes several hours to classify all the points.
+Possible areas of speedup are a better algorithm for finding the nearest neighbors or parallelizing the nearest neighbor search.
 

### 2013-05-24 Digit Recognition/Main.py
who knows

@@ -5,6 +5,7 @@ import Test
 import Train
 import sys
 import Kernels
+from multiprocessing.pool import Pool
 
 output = None
 train_input = None
@@ -53,13 +54,17 @@ def main():
 
     output = open("./data/output.txt", "w+")
 
-    knn = KNN.KNN(train_data, 1)
+
     classifications = list()
+    pool = Pool(processes=4)
+    results = list()
     for i in range(0, len(test_data)):
+        knn = KNN.KNN(train_data, 1)
         x_row = test_data[i, :]
-        classifications.append(knn.classifySample(x_row))
-        print "Correct Classification: %d Our Classification: %d" % (test_y[i], classifications[-1])
-        output.write(str(classifications[-1]) + "\n")
+        results.append(pool.apply(knn.classifySample, x_row))
+
+    for result in results:
+        print result.get()
 
     result = sum([np.sign(x) for x in np.subtract(test_y, classifications)])
     print result

### 2013-05-24 Digit Recognition/KNN.py
Made some updates and removed broken threading code. Final updated milestone report.

@@ -45,6 +45,7 @@ class KNN:
     # train_data is a matrix of where the first column holds the 
     # Y values and the other columns are features.
     def __init__(self, train_data, k=1, report=True):
+        print(train_data.shape)
         self.k = k
         self.train_data = train_data
         self.report = report

### 2013-05-24 Digit Recognition/KNNCrossValidation.py

@@ -4,7 +4,7 @@ import sys
 
 # Use cross validation to fine tune k
 # k_values is a list of values of k to consider
-def crossValidate(k_values, train_data):
+def crossValidate(k_values, num_folds, train_data):
     train = train_data
     errors = [0] * len(k_values)
 
@@ -13,17 +13,18 @@ def crossValidate(k_values, train_data):
         k = k_values[i]
 
         # Exclude k samples from our training set
-        for j in range(len(train_data)/k):
-            index = j * k
-            to_test = train[index:(index + k),:]
-            to_train = np.delete(train, range(index, index + k), axis=0)
+        fold_size = len(train_data) / num_folds
+        for j in range(num_folds):
+            index = j * fold_size
+            to_test = train[index:(index + fold_size), :]
+            to_train = np.delete(train, range(index, index + fold_size), axis=0)
             knn = KNN.KNN(to_train, k)
             
             for x in to_test:
                 x_sample = x[1:]
                 y = x[0]
                 if knn.classifySample(x_sample) != y:
-                    errors[i] += 1
+                    errors[i] += 1.0 / num_folds
     print errors
     return errors.index(min(errors))
 

### 2013-05-24 Digit Recognition/Main.py

@@ -64,20 +64,19 @@ def main():
     # KNN DRIVER
     #
     # Cross validation
-    k = cv.crossValidate([1,2,3,4,5,10,15,20], train_data[0:1000,)
+
+    k = cv.crossValidate([1,2,3,4,5,10,15,20], 10, train_data[0:10000])
     output = open("./data/output.txt", "w+")
     
     knn = KNN.KNN(train_data, k)
     classifications = list()
-    pool = Pool(processes=4)
     results = list()
     for i in range(0, len(test_data)):
         knn = KNN.KNN(train_data, 1)
         x_row = test_data[i, :]
-        results.append(pool.apply(knn.classifySample, x_row))
+        results.append(knn.classifySample, x_row)
 
-    for result in results:
-        print result.get()
+
 
     result = sum([np.sign(x) for x in np.subtract(test_y, classifications)])
     print result

### 2013-05-24 Writeup/milestone.tex
Moved writeup code into its own folder

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{hyperref}
\usepackage[margin=1.5in]{geometry}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

 \title{Digit Recognition Using KNN and SVM}

\author{
Harnoor Singh \\
Department of Computer Science \& Engineering\\
University of Washington \\
\texttt{hsingh@cs.washington.edu} \\
\And
Brian Walker \\
Department of Computer Science \& Engineering \\
University of Washington \\
\texttt{bdwalker@cs.washington.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

\section{Objective}

As described in our project proposal, we are using the Kaggle digit dataset to classify written numbers.

Our goal is to implement digit recognition using K Nearest Neighbors (KNN) and a Support Vector Machine (SVM) and compare the two approaches. A stretch objective is to enable data that isn't a part of the Kaggle dataset to work with our algorithms. More specifically, we would like to be able to classify some basic mathematical symbols in addition to digits.  This would allow us to be able to classify and evaluate simple mathematical expressions.  There is some challenges associated with this.  We would need to create our own data set as well as normalize the input to match with the Kaggle input.  Thus, this is a stretch goal that we hope to get to but only if time allows.

\section{KNN}

\subsection{KNN Progress}

We have made significant progress in both the KNN algorithm and the SVM algorithm. Our KNN algorithm is completely implemented k-fold cross validation to tune our parameter K.

The Kaggle dataset provides a larger training and testing set. However, we do not have the gold standard labels for the testing data. In order to evaluate the accuracy we have to submit our output into the Kaggle Digit Recognition competition.  Our implementation did surprisingly well, placing us at 263rd place out of 1237 with a successful classification rate of .97114.  This error rate was achieved with a 1NN implementation. Since then, we have reclassified and resubmitted with 15NN, which dropped our accuracy to .963.  We are quite happy with the results for such a simple algorithm. After performing our cross-validation on a small subset, it would appear that 1NN may be the best choice for K.  If this is the case then there is little optimization we can do to improve our implementation.  We are hoping the flexibility of a kerneled SVM will allow us to improve our classification.

As we discuss below, we will run cross-validation on the entire data set for the final report to ensure we have chosen the best candidate for K.  We will also run some more classification runs on the test data.  The current classification takes around 7.5 hours on my home computer which has 16Gb of RAM and an Intel Extreme Core I7 processor.  We will implement a threading procedure to speed up the classification before the final report to allow us to run more testing.

\subsection{KNN Procedure and Optimization}

To compute the K nearest neighbors, we take the euclidean difference between the feature we are classifying, $x_1$, and every feature in our training set.

\begin{center}
$\sum\limits_{i=1}^D {(x_{1i} - x_{2i})^2}$
\end{center}
Where D is the number of features for a sample.

Once we have calculated the K closest features, a simple voting method is used to classify $X_1$.

In order to select a value of K we do k-fold cross validation on a subset of the training set. The data below was sampled using 1000 samples.  We will perform cross validation with a larger data set for the final report but due to time constraints and the time required to do k-fold validation on the dataset provided by Kaggle, we limited the number of samples for this report.

We performed the k-fold validation by splitting the training data into k evenly divided sets.  We then use k-1 of these as "training" data and classify the remaining fold using the k-1 other sets.  We then rotate through the subsets and test with each one. We then average the results once testing has been done with each of the subsets.  We do this for different values of KNN and pick a value of K that gives us the best results.

The K value that led to the lowest error rate is used to build our final classifier. The result of our cross validation can be seen in Table 1 below.

\begin{table}[ht]
\caption{Cross Validation Selection of K} % title of Table
\centering  % used for centering table
\begin{tabular}{c c} 
\hline\hline
K & Misclassified \\ [1.0ex]
\hline 
1 & 135 \\
2 & 164 \\
3 & 137 \\
4 & 148 \\
5 & 144 \\
10 & 163 \\
15 & 176 \\
20 & 191 \\ [1ex]
\hline
\end{tabular}
\label{table:nonlin}
\end{table}

There is a large disconnect between our full classification error as obtained from Kaggle and the error rate above. This can be attributed to a much lower number of training samples.

\section{SVM}

\subsection{SVM Progress}

We have made good progress on our kerneled SVM implementation.  We are working on some issues with performing regularization and tuning the additional parameters we need for SVM.  We our currently capable of running the SVM with the digit recognition data but we are not confident in the results. We suspect we have a day or two worth of work to get the SVM completely running.

The implementation has been designed such that we can plug in different kernels for easy testing.  These kernels are also "curried" to allow them to be easily instantiated with their own tuning parameters for passing to the SVM.  Thus we can easily optimize our classifier by performing cross-validation with both different kernels and different parameters to the kernels themselves.  It is our hope that this flexibility will allow us to classify the trickier digits in the competition.

\subsection{SVM Procedure and Optimization}

The SVM implementations we have worked with previously have involved only two classes. Because we have to classify the digits 0-9, we need to create an SVM that can classify from a broader range.  The two primary methods for a multivariate SVM classifier is a one vs. one or one vs. all implementation.  There has been some debate about which is better and there does not appear to be a clear winner from the research we have done.  We have chosen to implement the one vs. one method for out multivariate SVM.

For one vs. one we must train a classifier for every pair of possible classifications.  Since SVM is a binary classifier we must consider every possible combination.  Thus, we must train $n choose 2$ SVMs which compare two numbers. For example, one SVM would classify a digit as either a 1 or a 2. Another would classify a digit as a 1 or a 3, and so on. We get a prediction from all of these classifiers and take a vote to see which digit was most likely based on our SVMs.

In our training step, every time we make a mistake we log it. Future predictions are made based on past mistakes.

\begin{center}
$Classification = sign(\sum\limits_{i \in Mistakes} {\alpha_i(Y_i * k(X_i, X)) + w_0})$
\end{center}

A classification is considered a mistake if the truth label $y$ multiplied by our prediction is less than 0.

The biggest challenge thus far has been deciding how to regularize our SVM without relying on the weights. Because we use the kernel trick, the weights are never explicitly stored.

\section{Future Work}

Our primary focus moving forward is to get our SVM implementation completed. Most of the structure is in place, however we need to decide on our method of regularization. Once we have the regularization done we can analyze the effectiveness of our SVM. We will implement cross-validation to tune the regularization as well as the kernel and the kernel dimensions.  The SVM provides much more flexibility through parameters than KNN does and thus cross-validation will be very important to ensure that our classifier is effective.

For KNN it might be beneficial to attempt to increase performance. Classifying each sample takes about 1 second, and with the testing dataset holding about 280,000 samples it takes several hours to classify all the points. Possible areas of speedup are a better algorithm for finding the nearest neighbors or parallelizing the nearest neighbor search.

If we have enough time we will then attempt at making our own dataset for classifying mathematical symbols.  We have been unable to find a dataset that provides ground truth labels to symbol training data. If we cannot find a suitable provider for the data we will have to make our own.  We suspect that this will not be very accurate since we will only be able to make a limited number of samples. However, it is something we would like to explore.

\begingroup
\renewcommand{\section}[2]{}%
%\renewcommand{\chapter}[2]{}% for other classes
\bibliographystyle{unsrt}
\bibliography{references}	
\endgroup

\end{document}

### 2013-06-07 Digit Recognition/Main.py
FUCK THIS. Perceptron finished.

@@ -72,7 +72,7 @@ def trainPerceptron(train_file, kernel, output_file, dim):
 # Loads the test data and model, unpacks the model and 
 # initializes the testing framework.
 def runPerceptron(test_file, kernel, output_file, dim, model_file):
-    testData = loadTestData(test_file)
+    testData, testY = loadTestData(test_file)
     tester = Test.TestModel()
     perceptron = Perceptron.Classifier()
     realKernel = None
@@ -86,7 +86,7 @@ def runPerceptron(test_file, kernel, output_file, dim, model_file):
     model = perceptron.unpackModel(model)
     classifications = tester.predictWithModel(model, testData, realKernel, True)
     output = open(output_file, "w+")
-    pickle.dump(classifications, output)
+    output.writelines(classifications)
 
 def printUsage():
 

### 2013-06-07 Digit Recognition/Perceptron.py

@@ -14,7 +14,11 @@ class Classifier:
         mistake_y = mistakes[:, 0]
         mistake_x = mistakes[:, 1:]
 
-        return 1 if (np.sum(mistake_y * np.apply_along_axis(kernel, 1, mistake_x, x))) > 0 else -1
+        #print "Length of x: ", len(x)
+        #print "Length of mistakes_x: ", len(mistake_x[0])
+
+        c =  1 if (np.sum(mistake_y * np.apply_along_axis(kernel, 1, mistake_x, x))) > 0 else -1
+        return c
 
     # Method to train the model.  It accepts as arguments and ndarray of training data,
     # an ndarray of their corresponding labels, the penalty constant, the kernel to use
@@ -46,27 +50,28 @@ class Classifier:
     # Method that accepts testing data as an ndarray and a trained model.  It will
     # make predictions for the test data with the trained model and return the predictions
     # in an ndarray.
-    def testWithModel(self, x, kernel, model):
-        mistakes = self.unpackModel(model)
+    def testWithModel(self, x, kernel, mistakes):
         return np.apply_along_axis(self.classify, 1, x, mistakes, kernel)
 
     # Simple helper method to unpack everything in the model.
     def unpackModel(self, model):
-        print model
+        mistakes = model["weights"]
+        print  mistakes[(3.0, 4.0)][1]
         return model["weights"]
 
     # Helper method to pack everything into a dictionary that will be needed
     # for testing with the trained model.  It will save all the mistakes needed
     # to make a prediction, the constant C needed for weighting and the kernel
     # used for training which should be identical to the one used for testing.
-    def packModel(self, weights):
-        return {"weights": weights}
+    def packModel(self, mistakes):
+        print mistakes[(3.0, 4.0)][1]
+        return {"weights": mistakes}
 
     # Trains the SVM model using the data passed in as x_data.  It will return
     # a trained model that can be used for making predictions on test data.  The
     # method accepts an integer representing the penalty to assign to the margin,
     # the kernel, if any, to be used to map the data to a higher dimension, a step size
-    # eta and a boolean to indicate if the average loss should be printed during training.
+    # eta and a boolean to indicate if the traverage loss should be printed during training.
     def trainModel(self, x_data, kernel=Kernels.defaultKernel(), report=False):
 
         # +1 for intercept
@@ -76,5 +81,6 @@ class Classifier:
         # add a column of ones for the intercept
[…]

### 2013-06-07 Digit Recognition/Test.py

@@ -12,17 +12,28 @@ class TestModel:
         for i in range(predictions.shape[0]):
             lables[which.predictions[i,:]]
 
+    # Passed an unpacked model_dict
     def predictWithModel(self, model_dict, test_data, kernel, report=False):
         perceptron = Perceptron.Classifier()
-        votes = np.ndarray(shape=(test_data.shape[0], 10), dtype = int)
+        
+        votes = np.zeros(shape=(test_data.shape[0], 10), dtype = int)
+
         for couple in model_dict:
             predictions = perceptron.testWithModel(test_data, kernel, model_dict[couple])
+            print predictions
             for i in range(predictions.shape[0]):
                 if predictions[i] == 1:
                     votes[i, couple[0]] += 1
                 else:
                     votes[i, couple[1]] += 1
-        return predictions
+
+        classifications = list()
+
+        for i in range(len(votes)):
+            classification = str(votes[i].argmax()) + "\n"
+            print "Classified ", classification
+            classifications.append(classification)
+        return classifications
 
     def calculateLoss(self, predictions, labels):
         error = 0

### 2013-06-07 Digit Recognition/Train.py

@@ -88,13 +88,13 @@ class Trainer:
         trained_model = dict()
         for couple in self.processed_data.keys():
             arg = (self.processed_data[couple], kernel, report, plot)
-            #trained_model[couple] = self._trainSingleClass(self.processed_data[couple], kernel, report, plot)
             result = pool.apply_async(self._trainSingleClass, arg)
             threads[couple] = result
 
         trained_model = dict()
         for r in threads.keys():
             trained_model[r] = threads[r].get()
+            #Stored length 785
 
         perceptron = Perceptron.Classifier()
         model = perceptron.packModel(trained_model)

### 2013-06-09 Paper/paper.tex
Started writeup in Paper/*

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{hyperref}
\usepackage[margin=1.5in]{geometry}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

 \title{Digit Recognition: Nearest Neighbors vs. Perceptron}

\author{
Harnoor Singh \\
Department of Computer Science \& Engineering\\
University of Washington \\
\texttt{hsingh@cs.washington.edu} \\
\And
Brian Walker \\
Department of Computer Science \& Engineering \\
University of Washington \\
\texttt{bdwalker@cs.washington.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

\begin{abstract}

Optical Character Recognition (OCR) is a complex computer science problem which
has several effective machine learning solutions. We implemented the K Nearest
Neighbors (KNN) and Kernelized Perceptron algorithms and compared the
effectiveness of both methods in classifying a testing data set of 28,000
images. Each image is formatted as an array of 768 pixel values, with each value
indicating how dark each pixel was in the image. Both of our algorithms were
trained on a training data set of 42,000 images. We used cross validation to
tune the K value in KNN and to choose a kernel function and dimension value for
Perceptron. Despite its relative simplicity, we found that KNN was more
effective than Perceptron in classifying the digits. The nearest neighbors
algorithm correctly classified 96.857\% of the testing samples on a weighted
neighbors algorithm which looked at 4 neighbors. Our best Perceptron
implementation correctly classified 95.67\% of the testing samples using an
exponential kernel with a $\sigma = 5$. Despite its higher accuracy, the runtime
of KNN was 8x slower than that of the Perceptron.

\end{abstract}

\section{The Challenge}

Optical Character Recognition (OCR) is a computational problem where a machine
attempts to classify an image of text. We tackled a subset of this problem,
attempting to classify the digits 0-9 from data provided by the Kaggle Digit
Recognition
competition\footnote{\url{http://www.kaggle.com/c/digit-recognizer}}.

The Kaggle data provides two groups of data, 42,000 training samples with labels
and 28,000 unlabeled testing samples.

\begin{table}[ht]
\caption{Subset of Kaggle Data} % title
\centering 
\begin{tabular}{c c c c }

Label & Pixel 77 & Pixel 78 & Pixel 79 \\

8 & 0 & 0 & 0 \\
6 & 89 & 208 & 135 \\ [1ex]

\end{tabular}
\label{table:nonlin}
\end{table}

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
preferred typeface throughout. Paragraphs are separated by 1/2~line space,
with no indentation.

Paper title is 17~point, initial caps/lower case, bold, centered between
2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
thick. Allow 1/4~inch space above and below title to rules. All pages should
start at 1~inch (6~picas) from the top of the page.

%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.

For the final version, authors' names are
set in boldface, and each name is centered above the corresponding
address. The lead author's name is to be listed first (left-most), and
the co-authors' names (if different address) are set to follow. If
there is only one co-author, list both author and co-author side by side.

Please pay special attention to the instructions in section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

First level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 12. One line space before the first level
heading and 1/2~line space after the first level heading.

\subsection{Headings: second level}

Second level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the second level
heading and 1/2~line space after the second level heading.

\subsubsection{Headings: third level}

Third level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the third level
heading and 1/2~line space after the third level heading.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone, regardless of the formatter being used.

\subsection{Citations within the text}

Citations within the text should be numbered consecutively. The corresponding
number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
corresponding references are to be listed in the same order at the end of the
paper, in the \textbf{References} section. (Note: the standard
\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
references themselves, any style is acceptable as long as it is used
consistently.

As submission is double blind, refer to your own published work in the 
third person. That is, use ``In the previous work of Jones et al.\ [4]'',
not ``In our previous work [4]''. If you cite your other papers that
are not widely available (e.g.\ a journal paper under review), use
anonymous author names in the citation, e.g.\ an author of the
form ``A.\ Anonymous''.

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first footnote} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches
(12~picas).\footnote{Sample of the second footnote}

\subsection{Figures}

All artwork must be neat, clean, and legible. Lines should be dark
enough for purposes of reproduction; art work should not be
hand-drawn. The figure number and caption always appear after the
figure. Place one line space before the figure caption, and one line
space after the figure. The figure caption is lower case (except for
first word and proper nouns); figures are numbered consecutively.

Make sure the figure caption does not get separated from the figure.
Leave sufficient space to avoid splitting the figure and figure caption.

You may use color figures. 
However, it is best for the
figure captions and the paper body to make sense if the paper is printed
either in black/white or in color.
\begin{figure}[h]
\begin{center}
%\framebox[4.0in]{$\;$}
\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

\subsection{Tables}

All tables must be centered, neat, clean and legible. Do not use hand-drawn
tables. The table number and title always appear before the table. See
Table~\ref{sample-table}.

Place one line space before the table title, one line space after the table
title, and one line space after the table. The table title must be lower case
(except for first word and proper nouns); tables are numbered consecutively.

\begin{table}[t]
\caption{Sample table title}
\label{sample-table}
\begin{center}
\begin{tabular}{ll}
\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
\\ \hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\section{Final instructions}
Do not change any aspects of the formatting parameters in the style files.
In particular, do not modify the width or length of the rectangle the text
should fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PostScript or PDF files}

Please prepare PostScript or PDF files with paper size ``US Letter'', and
not, for example, ``A4''. The -t
letter option on dvips will produce US Letter files.

Fonts were the main cause of problems in the past years. Your PDF file must
only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
to achieve this.

\begin{itemize}

\item You can check which fonts a PDF files uses.  In Acrobat Reader,
select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts
are also acceptable for NIPS. Please see
\url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item LaTeX users:

\begin{itemize}

\item Consider directly generating PDF files using \verb+pdflatex+
(especially if you are a MiKTeX user). 
PDF figures must be substituted for EPS figures, however.

\item Otherwise, please generate your PostScript and PDF files with the following commands:
\begin{verbatim} 
dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
ps2pdf mypaper.ps mypaper.pdf
\end{verbatim}

Check that the PDF files only contains Type 1 fonts. 
%For the final version, please send us both the Postscript file and
%the PDF file.

\item xfig "patterned" shapes are implemented with 
bitmap fonts.  Use "solid" shapes instead. 
\item The \verb+\bbold+ package almost always uses bitmap
fonts.  You can try the equivalent AMS Fonts with command
\begin{verbatim}
\usepackage[psamsfonts]{amssymb}
\end{verbatim}
 or use the following workaround for reals, natural and complex: 
\begin{verbatim}
\newcommand{\RR}{I\!\!R} %real numbers
\newcommand{\Nat}{I\!\!N} %natural numbers 
\newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}

\item Sometimes the problematic fonts are used in figures
included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
way to clean such figures. For black and white figures, slightly better
results can be achieved with program \verb+potrace+.
\end{itemize}
\item MSWord and Windows users (via PDF file):
\begin{itemize}
\item Install the Microsoft Save as PDF Office 2007 Add-in from
\url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
\item Select ``Save or Publish to PDF'' from the Office or File menu
\end{itemize}
\item MSWord and Mac OS X users (via PDF file):
\begin{itemize}
\item From the print menu, click the PDF drop-down box, and select ``Save
as PDF...''
\end{itemize}
\item MSWord and Windows users (via PS file):
\begin{itemize}
\item To create a new printer
on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
\url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
AdobePS driver for it to take effect.
\item To produce the ps file, select ``Print'' from the MS app, choose
the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
\item Set ``TrueType Font'' to be ``Download as Softfont''
\item Open the ``PostScript Options'' folder
\item Select ``PostScript Output Option'' to be ``Optimize for Portability''
\item Select ``TrueType Font Download Option'' to be ``Outline''
\item Select ``Send PostScript Error Handler'' to be ``No''
\item Click ``OK'' three times, print your file.
\item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
the PS file. In Acrobat, check the option ``Embed all fonts'' if
applicable.
\end{itemize}

\end{itemize}
If your file contains Type 3 fonts or non embedded TrueType fonts, we will
ask you to fix it.

\subsection{Margins in LaTeX}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+
from the graphicx package. Always specify the figure width as a multiple of
the line width as in the example below using .eps graphics
\begin{verbatim}
   \usepackage[dvips]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.eps} 
\end{verbatim}
or % Apr 2009 addition
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.pdf} 
\end{verbatim}
for .pdf graphics. 
See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})

A number of width problems arise when LaTeX cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All
acknowledgments go at the end of the paper. Do not include 
acknowledgments in the anonymized submission, only in the 
final paper. 
\subsubsection*{References}

References follow the acknowledgments. Use unnumbered third level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to `small' (9-point) 
when listing the references. {\bf Remember that this year you can use
a ninth page as long as it contains \emph{only} cited references.}

\small{
[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
and T.K. Leen (eds.), {\it Advances in Neural Information Processing
Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.

[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
Realistic Neural Models with the GEneral NEural SImulation System.}
New York: TELOS/Springer-Verlag.

[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
and recall at excitatory recurrent synapses and cholinergic modulation
in rat hippocampal region CA3. {\it Journal of Neuroscience}
{\bf 15}(7):5249-5262.
}

\end{document}

### 2013-06-09 Paper/paper.tex
Update writeup code

@@
 
-\section{The Challenge}
+\section{The Problem}
 
+\subsection{Image Data}
 Optical Character Recognition (OCR) is a computational problem where a machine
@@
 The Kaggle data provides two groups of data, 42,000 training samples with labels
-and 28,000 unlabeled testing samples.
+and 28,000 unlabeled testing samples. Each training sample contains a 48 pixel
+by 48 pixel image containing a handwritten number. The image is represented as
+an array of 768 pixel values which represent how dark each pixel of the image
+is. Table 1 is an example of a training sample.
 
 \begin{table}[ht]
-\caption{Subset of Kaggle Data} % title
+\textbf{\caption{Subset of Kaggle Data}} % title
 \centering 
-\begin{tabular}{c c c c }
-
+\begin{tabular}{c | c c c }
 Label & Pixel 77 & Pixel 78 & Pixel 79 \\
-
+\hline
 8 & 0 & 0 & 0 \\
 6 & 89 & 208 & 135 \\ [1ex]
-
 \end{tabular}
@@
 
-\section{General formatting instructions}
-\label{gen_inst}
+\section{K Nearest Neighbors (KNN)}
 
-The text must be confined within a rectangle 5.5~inches (33~picas) wide and
-9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
-Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
-preferred typeface throughout. Paragraphs are separated by 1/2~line space,
-with no indentation.
+\subsection{Overview}
 
-Paper title is 17~point, initial caps/lower case, bold, centered between
-2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
-thick. Allow 1/4~inch space above and below title to rules. All pages should
-start at 1~inch (6~picas) from the top of the page.
+K Nearest Neighbors is a relatively simple algorithm which often gets surprisingly
+good results given its simplicity. Given a testing sample $X$, the algorithm
+looks through all the training samples it has previously seen and finds the
+\textit{K} most similar samples. The algorithm then implements some form of
+voting mechanism among the \textit{K} samples to classify the testing sample as
+a digit 0-9. The voting mechanism can either be as straightforward as selecting the
+majority label, to something more complicated which involves weighting the
+samples based on some distance metric.
 
-%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.
+\subsection{Implementation Details}
 
-For the final version, authors' names are
-set in boldface, and each name is centered above the corresponding
-address. The lead author's name is to be listed first (left-most), and
-the co-authors' names (if different address) are set to follow. If
-there is only one co-author, list both author and co-author side by side.
+The first step in implementing KNN is to find the \textit{K} nearest
+neighbors. This is accomplished by taking the euclidean distance between the
+testing sample we are classifying and every training sample in our training data
+set.
 
-Please pay special attention to the instructions in section \ref{others}
-regarding figures, tables, acknowledgments, and references.
+\begin{center}
+$Euclidean(TestX, TrainX) = \sum\limits_{i \in TrainX} {(TestX - TrainX_i)^2}$
+\end{center}
 
-\section{Headings: first level}
-\label{headings}
+Because the image data is defined as an array of 768 pixels, $TestX - TrainX_i$
+is actually doing a 768 dimension comparison between data points.
 
-First level headings are lower case (except for first word and proper nouns),
-flush left, bold and in point size 12. One line space before the first level
-heading and 1/2~line space after the first level heading.
+Once we identified the \textit{K} nearest neighbors to our testing sample, we
+had to use this information to classify the testing point. We considered two
+approaches when deciding how to perform this classification.
 
-\subsection{Headings: second level}
+\subsubsection{Unweighted KNN}
 
-Second level headings are lower case (except for first word and proper nouns),
-flush left, bold and in point size 10. One line space before the second level
-heading and 1/2~line space after the second level heading.
+The first method we implemented to classify our testing sample was a simple
+majority voting algorithm which selects the majority label from the \textit{K}
+closest neighbors. For example, if we ran the nearest neighbors algorithm with
+$K = 5$ and 3 of the neighbors had a label of 6, we would classify our testing
+sample as a 6.
 
-\subsubsection{Headings: third level}
+\subsubsection{Weighted KNN}
 
-Third level headings are lower case (except for first word and proper nouns),
-flush left, bold and in point size 10. One line space before the third level
-heading and 1/2~line space after the third level heading.
+Given the \textit{K} nearest neighbors, the weighted version of the algorithm
+multiplies each label by a weight which indicates how close it is to our testing
+sample. These weights help to emphasize points that are closer to our testing
+sample while reducing the impact further away training points have on
+classifying our testing sample.
+
+\subsection{Cross Validation}
+
+Our implementation of KNN has a number of parameters which require
+tuning. In particular, we need to decide whether or not to use the weighted
+version of our KNN algorithm and how many neighbors to consider when
+implementing our algorithm.
+
+To do this, we implemented K folds cross validation using 5,000 training
+samples. The reduced training set allowed us to run our algorithm using various
+configurations to see which configuration was most accurate.
+
+\begin{table}[ht]
+\textbf{\caption{KNN Cross Validation Error Rates}} % title
+\centering
+\begin{tabular}{c | c c }
+K & Unweighted KNN Error & Weighted KNN Error \\
+\hline
+1 & .077 & .077 \\
+2 & .092 & .092 \\
+3 & .075 & .075 \\
+4 & .078 & .073 \\
+5 & .078 & .075 \\
+10 & .085 & .079 \\
+15 & .095 & .089 \\
+20 & .102 & .098 \\ [1ex]
+\end{tabular}
+\label{table:nonlin}
+\end{table}
+
+Our cross validation results suggest that running weighted KNN with $K = 4$
+yields the most successful classification rate. One caveat to this is that our
+cross validation code ran using 5,000 training samples instead of the 42,000
+samples in our full training set. It is possible that data we used to cross
+validate is no representative of the full data set.
+
+\subsection{Runtime}
+
+\subsection{Results}
+
+\section{Kernelized Perceptron}
+\subsubsection{Overview}
+
+\subsubsection{Implementation Details}
+
+\subsubsection{Results}
 

### 2013-06-10 Paper/paper.tex
First draft of the paper done

@@
 \usepackage[margin=1.5in]{geometry}
+\usepackage{graphicx}
+
 %\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09
@@
 \end{tabular}
-\label{table:nonlin}
+\label{table:Data}
 \end{table}
@@
 
+\begin{center}
+\includegraphics[width=406px]{KNN_CrossValidation.png}
+\end{center}
+
 Our cross validation results suggest that running weighted KNN with $K = 4$
@@
 
-\subsection{Runtime}
+\subsection{Results}
+
+Our KNN implementation was extremely successful, correctly classifying 96.857\%
+of the testing samples. One downside to it however is the runtime. Training is
+near instantaneous since the algorithm simply loads the data into an array. To
+classify a sample the algorithm needs to compute the euclidean distance between
+the testing sample and every training sample. In the case of the full data set
+this is 42,000 samples, each of which has 768 dimensions. As a result, each
+testing sample takes approximately. Altogether it takes close to 8 hours to
+classify all 28,000 testing samples.
+
+\section{Kernelized Perceptron}
+\subsection{Overview}
+
+The perceptron algorithm attempts to separate a training data set into two
+classes and attempts to classify future testing samples as belonging to one of
+these two classes. It does this by storing a list of all the mistakes it has
+made previously and using this knowledge to adjust future predictions in the
+hopes of improving classification accuracy.
+
+In its most basic implementation, the perceptron algorithm divides the data
+linearly. However, the data is often not linearly separable. To address this
+issue we use a kernel function to create a non-linear decision boundary. This
+allows us to have a more complex decision boundary which potentially fits the
+data better.
+
+\subsection{Implementation Details}
+
+Before we can begin classifying test points, we need to train the perceptron
+algorithm using our training data set. The training process attempts
+to classify each training point sequentially. If the perceptron correctly
+classifies the point nothing changes and the training process moves on to the
+next training sample. If the training sample is incorrectly classified, however,
+the mistake is added to an array of mistakes which is taken into consideration
+for future classifications.
+
+One problem that needs to be addressed is that a standard perceptron is a binary
+classifier meaning it selects between one of two classes. We resolved this issue
+by training $10 \choose 2 = 45$ classifiers. Each classifier is responsible for
+classifying between two digits. For example, we would train a perceptron to
+classify between 0 and 1, and a second to classify between 0 and 2, and so
+on. When we classify we vote between all of our trained classifiers and output
+the majority element.
+
+The actual classification is performed by:
+
+\begin{center}
+$classify(X) = \sum\limits_{i \in Mistakes}{Y_iK(X_i, X)}$
+\end{center}
+
+The kernel function is represented by $K(X_i, X)$. There are a number of choices
+for this kernel function.
+
+\subsection{Cross Validation for Kernel Selection}
+
+We implemented two kernel functions to use with our perceptron algorithm. One is
+a polynomial kernel which takes the form:
+
+\begin{center}
+$Polynomial(x, y, d) = (dot(x, y) + 1)^d$
+\end{center}
+
+The second is an exponential kernel:
+
+\begin{center}
+$Exponential(x, y, \sigma) = exp(-\frac{||x - y||}{2\sigma^2})$
+\end{center}
+
+Both of these kernels have a parameter which needs to be tuned, $d$ in the case of
+the polynomial kernel and $\sigma$ in the case of the exponential kernel. We
+used cross validation to find out which kernel gave us the best result as well
+as which value of $d$ and $\sigma$ was optimal.
+
+\begin{table}[ht]
+\textbf{\caption{Perceptron Cross Validation Success Rates}} % title
+\centering
+\begin{tabular}{c | c c }
+Kernel & $d, \sigma = 5$ & $d, \sigma = 10$ \\
+\hline
+Polynomial & 95.67\% & 94.52\% \\
+Exponential & 93.84\% & 96.3\% \\ [1ex]
+\end{tabular}
+\label{table:perceptron}
+\end{table}
+
+\begin{center}
+\includegraphics[width=406px]{Perceptron_CrossValidation.png}
+\end{center}
 
@@
 
-\section{Kernelized Perceptron}
-\subsubsection{Overview}
+Our cross validation indicates that the exponential kernel with a $\sigma$ of 10
+leads to the highest success rate of 96.3\%. An interesting observation is that
+our polynomial kernel actually became less successful when it's dimension was
+increased whereas the exponential kernel got better. We suspect that this is
+because the polynomial kernel began to overfit the training data.
 
-\subsubsection{Implementation Details}
+Finally, the runtime of the perceptron algorithm is a little over an hour. It
+takes approximately 45 minutes to train and about 20 minutes to test the
+data. The faster runtime allowed us to cross validate on the full data set.
 
-\subsubsection{Results}
+\section{Conclusion}
+Despite its seeming simplicity, we found that KNN was extremely successful
+when classifying data points in the Kaggle set.
 
-\section{Citations, figures, tables, references}
-\label{others}
-
-These instructions apply to everyone, regardless of the formatter being used.
-
-\subsection{Citations within the text}
-
-Citations within the text should be numbered consecutively. The corresponding
-number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
-corresponding references are to be listed in the same order at the end of the
-paper, in the \textbf{References} section. (Note: the standard
-\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
-references themselves, any style is acceptable as long as it is used
-consistently.
-
-As submission is double blind, refer to your own published work in the 
-third person. That is, use ``In the previous work of Jones et al.\ [4]'',
-not ``In our previous work [4]''. If you cite your other papers that
-are not widely available (e.g.\ a journal paper under review), use
-anonymous author names in the citation, e.g.\ an author of the
-form ``A.\ Anonymous''.
-
-\subsection{Footnotes}
-
-Indicate footnotes with a number\footnote{Sample of the first footnote} in the
-text. Place the footnotes at the bottom of the page on which they appear.
-Precede the footnote with a horizontal rule of 2~inches
-(12~picas).\footnote{Sample of the second footnote}
-
-\subsection{Figures}
-
-All artwork must be neat, clean, and legible. Lines should be dark
-enough for purposes of reproduction; art work should not be
-hand-drawn. The figure number and caption always appear after the
-figure. Place one line space before the figure caption, and one line
-space after the figure. The figure caption is lower case (except for
-first word and proper nouns); figures are numbered consecutively.
-
-Make sure the figure caption does not get separated from the figure.
-Leave sufficient space to avoid splitting the figure and figure caption.
-
-You may use color figures. 
-However, it is best for the
-figure captions and the paper body to make sense if the paper is printed
-either in black/white or in color.
-\begin{figure}[h]
 \begin{center}
-%\framebox[4.0in]{$\;$}
-\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-\end{center}
-\caption{Sample figure caption.}
-\end{figure}
-
-\subsection{Tables}
-
-All tables must be centered, neat, clean and legible. Do not use hand-drawn
-tables. The table number and title always appear before the table. See
-Table~\ref{sample-table}.
-
-Place one line space before the table title, one line space after the table
-title, and one line space after the table. The table title must be lower case
-(except for first word and proper nouns); tables are numbered consecutively.
-
-\begin{table}[t]
-\caption{Sample table title}
-\label{sample-table}
-\begin{center}
-\begin{tabular}{ll}
-\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
-\\ \hline \\
-Dendrite         &Input terminal \\
-Axon             &Output terminal \\
-Soma             &Cell body (contains cell nucleus) \\
+\begin{tabular}{ c| c c}
+	\hline
+Algorithm & Percent Success & runtime\\
+	\hline
+Nearest Neighbors, $k = 4$   & 96.857\% & $\sim$8 hours\\
+Perceptron   & 96.300\%  & $\sim$1 hour\\
 \end{tabular}
 \end{center}
-\end{table}
 
-\section{Final instructions}
-Do not change any aspects of the formatting parameters in the style files.
-In particular, do not modify the width or length of the rectangle the text
-should fit into, and do not change font sizes (except perhaps in the
-\textbf{References} section; see below). Please note that pages should be
-numbered.
+Table 4 summarizes the accuracy and runtimes of weighted KNN with $K = 4$ and
+perceptron with the polynomial kernel and $d = 10$. KNN is clearly more
+accurate, however because perceptron is 8x faster and almost as accurate, there
+could be many cases where it is a better choice.
 
-\section{Preparing PostScript or PDF files}
+In addition to being faster, perceptron likely has a lower memory
+footprint. While KNN must store every sample in the training set, perceptron
+only stores mistakes which will be a subset of the entire training set.
 
-Please prepare PostScript or PDF files with paper size ``US Letter'', and
-not, for example, ``A4''. The -t
-letter option on dvips will produce US Letter files.
-
-Fonts were the main cause of problems in the past years. Your PDF file must
-only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
-to achieve this.
+\section{Future Work}
+There are three specific expansions to our work we would have liked to explore
+had we had time.
 
 \begin{itemize}
-
-\item You can check which fonts a PDF files uses.  In Acrobat Reader,
-select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-available out-of-the-box on most Linux machines.
-
-\item The IEEE has recommendations for generating PDF files whose fonts
-are also acceptable for NIPS. Please see
-\url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
-
-\item LaTeX users:
-
-\begin{itemize}
-
-\item Consider directly generating PDF files using \verb+pdflatex+
-(especially if you are a MiKTeX user). 
-PDF figures must be substituted for EPS figures, however.
-
-\item Otherwise, please generate your PostScript and PDF files with the following commands:
-\begin{verbatim} 
-dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
-ps2pdf mypaper.ps mypaper.pdf
-\end{verbatim}
-
-Check that the PDF files only contains Type 1 fonts. 
-%For the final version, please send us both the Postscript file and
-%the PDF file.
-
-\item xfig "patterned" shapes are implemented with 
-bitmap fonts.  Use "solid" shapes instead. 
-\item The \verb+\bbold+ package almost always uses bitmap
-fonts.  You can try the equivalent AMS Fonts with command
-\begin{verbatim}
-\usepackage[psamsfonts]{amssymb}
-\end{verbatim}
- or use the following workaround for reals, natural and complex: 
-\begin{verbatim}
-\newcommand{\RR}{I\!\!R} %real numbers
-\newcommand{\Nat}{I\!\!N} %natural numbers 
-\newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-
-\item Sometimes the problematic fonts are used in figures
-included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
-way to clean such figures. For black and white figures, slightly better
-results can be achieved with program \verb+potrace+.
-\end{itemize}
-\item MSWord and Windows users (via PDF file):
-\begin{itemize}
-\item Install the Microsoft Save as PDF Office 2007 Add-in from
-\url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
-\item Select ``Save or Publish to PDF'' from the Office or File menu
-\end{itemize}
-\item MSWord and Mac OS X users (via PDF file):
-\begin{itemize}
-\item From the print menu, click the PDF drop-down box, and select ``Save
-as PDF...''
-\end{itemize}
-\item MSWord and Windows users (via PS file):
-\begin{itemize}
-\item To create a new printer
-on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
-\url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
-AdobePS driver for it to take effect.
-\item To produce the ps file, select ``Print'' from the MS app, choose
-the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
-\item Set ``TrueType Font'' to be ``Download as Softfont''
-\item Open the ``PostScript Options'' folder
-\item Select ``PostScript Output Option'' to be ``Optimize for Portability''
-\item Select ``TrueType Font Download Option'' to be ``Outline''
-\item Select ``Send PostScript Error Handler'' to be ``No''
-\item Click ``OK'' three times, print your file.
-\item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
-the PS file. In Acrobat, check the option ``Embed all fonts'' if
-applicable.
+\item Expanded cross validation for KNN using the full training set. This would
+  help validate that the weighted voting algorithm with $K = 4$ leads to the
+  most accurate. 
+\item More kernels and a greater assortment of $\sigma$ and $d$ values to
+  see if we can increase the effectiveness of the perceptron algorithm. 
+\item An SVM implementation which regularizes the data to avoid overfitting. If
+  the perceptron implementation overfits the data, this would be a potential
+  way to increase the accuracy.
 \end{itemize}
 
-\end{itemize}
-If your file contains Type 3 fonts or non embedded TrueType fonts, we will
-ask you to fix it.
-
-\subsection{Margins in LaTeX}
-
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+
-from the graphicx package. Always specify the figure width as a multiple of
-the line width as in the example below using .eps graphics
-\begin{verbatim}
-   \usepackage[dvips]{graphicx} ... 
-   \includegraphics[width=0.8\linewidth]{myfile.eps} 
-\end{verbatim}
-or % Apr 2009 addition
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ... 
-   \includegraphics[width=0.8\linewidth]{myfile.pdf} 
-\end{verbatim}
-for .pdf graphics. 
-See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})
-
-A number of width problems arise when LaTeX cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command.
-
-\subsubsection*{Acknowledgments}
-
-Use unnumbered third level headings for the acknowledgments. All
-acknowledgments go at the end of the paper. Do not include 
-acknowledgments in the anonymized submission, only in the 
-final paper. 
-\subsubsection*{References}
-
-References follow the acknowledgments. Use unnumbered third level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to `small' (9-point) 
-when listing the references. {\bf Remember that this year you can use
-a ninth page as long as it contains \emph{only} cited references.}
-
-\small{
-[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
-for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
-and T.K. Leen (eds.), {\it Advances in Neural Information Processing
-Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
-Realistic Neural Models with the GEneral NEural SImulation System.}
-New York: TELOS/Springer-Verlag.
-
-[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
-and recall at excitatory recurrent synapses and cholinergic modulation
-in rat hippocampal region CA3. {\it Journal of Neuroscience}
-{\bf 15}(7):5249-5262.
-}
+\begingroup
+\renewcommand{\section}[2]{}%
+%\renewcommand{\chapter}[2]{}% for other classes
+\bibliographystyle{unsrt}
+\bibliography{references}	
+\endgroup
 

### 2013-06-10 Paper/paper.tex
Done!

@@
 neighbors algorithm which looked at 4 neighbors. Our best Perceptron
-implementation correctly classified 95.67\% of the testing samples using an
-exponential kernel with a $\sigma = 5$. Despite its higher accuracy, the runtime
+implementation correctly classifies 96.3\% of the testing samples using an
+exponential kernel with a $\sigma = 10$. Despite its higher accuracy, the runtime
 of KNN was 8x slower than that of the Perceptron.
@@
 an array of 768 pixel values which represent how dark each pixel of the image
-is. Table 1 is an example of a training sample.
+is. Table 1 is an example of a training data sample.
 
@@
 
-K Nearest Neighbors is a relatively simple algorithm which often gets surprisingly
-good results given its simplicity. Given a testing sample $X$, the algorithm
-looks through all the training samples it has previously seen and finds the
-\textit{K} most similar samples. The algorithm then implements some form of
-voting mechanism among the \textit{K} samples to classify the testing sample as
-a digit 0-9. The voting mechanism can either be as straightforward as selecting the
-majority label, to something more complicated which involves weighting the
-samples based on some distance metric.
+K Nearest Neighbors is a relatively simple algorithm which often gets
+surprisingly good results given its simplicity. Given a testing sample $X$, the
+algorithm looks through all the training samples it has previously seen and
+finds the \textit{K} most similar samples, known as the nearest neighbors. The
+algorithm then implements some form of a voting mechanism among the \textit{K}
+samples to classify the testing sample as a digit 0-9. The voting mechanism can
+be as straightforward as selecting the majority label, or it can be more
+complex, involving weighting samples based on some distance metric.
 
@@
 
+The weight function is an inverse distance function of the form:
+
+\begin{center}
+$Weight(TrainY, TrainX, TestX) = \frac{1}{Euclidean(TestX, TrainX)}*TrainY$
+\end{center}
+
 \subsection{Cross Validation}
@@
 \hline
-1 & .077 & .077 \\
-2 & .092 & .092 \\
-3 & .075 & .075 \\
-4 & .078 & .073 \\
-5 & .078 & .075 \\
-10 & .085 & .079 \\
-15 & .095 & .089 \\
-20 & .102 & .098 \\ [1ex]
+1 & 7.7\% & 7.7\% \\
+2 & 9.2\% & 9.2\% \\
+3 & 7.5\% & 7.5\% \\
+4 & 7.8\% & 7.3\% \\
+5 & 7.8\% & 7.5\% \\
+10 & 8.5\% & 7.9\% \\
+15 & 9.5\% & 8.9\% \\
+20 & 10.2\% & 9.8\% \\ [1ex]
 \end{tabular}
@@
 samples in our full training set. It is possible that data we used to cross
-validate is no representative of the full data set.
+validate is not representative of the full data set.
 
@@
 Our KNN implementation was extremely successful, correctly classifying 96.857\%
-of the testing samples. One downside to it however is the runtime. Training is
+of the testing samples. One downside to it, however, is the runtime. Training is
 near instantaneous since the algorithm simply loads the data into an array. To
@@
 this is 42,000 samples, each of which has 768 dimensions. As a result, each
-testing sample takes approximately. Altogether it takes close to 8 hours to
-classify all 28,000 testing samples.
+testing sample takes approximately one second to run. Altogether it takes close
+to 8 hours to classify all 28,000 testing samples.
 
@@
 classifier meaning it selects between one of two classes. We resolved this issue
-by training $10 \choose 2 = 45$ classifiers. Each classifier is responsible for
+by training $10 \choose 2$ $=45$ classifiers. Each classifier is responsible for
 classifying between two digits. For example, we would train a perceptron to
@@
 The kernel function is represented by $K(X_i, X)$. There are a number of choices
-for this kernel function.
+for this kernel function which we will explore in the next section.
 
@@
 takes approximately 45 minutes to train and about 20 minutes to test the
-data. The faster runtime allowed us to cross validate on the full data set.
+data. The faster runtime allowed us to cross validate on the full data set so
+our cross validation results are more accurate than those we obtained for KNN.
 
@@
 
-\begin{center}
+\begin{table}[ht]
+\textbf{\caption{KNN vs. Perceptron on Full Testing Set}}
+\centering
 \begin{tabular}{ c| c c}
@@
 \end{tabular}
-\end{center}
+\end{table}
 
@@
 perceptron with the polynomial kernel and $d = 10$. KNN is clearly more
-accurate, however because perceptron is 8x faster and almost as accurate, there
-could be many cases where it is a better choice.
+accurate, however because perceptron is nearly 8x faster, there could be many
+cases where it is a better choice.
 
@@
 footprint. While KNN must store every sample in the training set, perceptron
-only stores mistakes which will be a subset of the entire training set.
+only stores mistakes which will be a subset of the training set.
+
+One reason why KNN may have been so successful in this trial is due to the large
+training data set and small number of classes to classify. With only 10 posible
+classifications, there are likely around 4,2000 training samples of each
+number. The odds of at least a few of those being similar to the test image
+intuitively seem high.
+
+An explanation for the perceptron algorithm being less successful than KNN is
+that it may have overfit our training data because there was no
+regularization. This could explain why our polynomial kernel did worse with a
+higher dimension kernel function.
 
@@
   help validate that the weighted voting algorithm with $K = 4$ leads to the
-  most accurate. 
+  most accurate classification. 
 \item More kernels and a greater assortment of $\sigma$ and $d$ values to

### 2013-06-11 Paper/paper.tex
paper.tex

@@
 $K = 5$ and 3 of the neighbors had a label of 6, we would classify our testing
-sample as a 6.
+sample as a 6. If there is a tie between two labels then the label is picked 
+arbitrarily.
 
@@
 
-The weight function is an inverse distance function of the form:
+Many different weight functions are available but we chose to use the inverse
+distance. This allows samples that are closest to one another to have the 
+largest impact while discounting those that are farthest away. The distance
+formula we used is below:
 
@@
 
+\subsection{Analyzation}
+
+The KNN algorithm is often very accurate but can be too slow for some applications. 
+In the case of OCR it proved to be the easiest to implement and the most accurate. 
+The algorithm uses a brute force tactic to compare a point against all other training
+points.  This allows a decision to be made with all the information available.  However,
+we were still only able to obtain a max of 96.857\% accuracy.  There are several possible
+reasons for this.  First, OCR is very difficult.  There are likely some hand written samples
+that simply do not have a good representation of nearest neighbors. Peoples handwriting 
+styles are different enough that some of the more unique interpretations of numbers did 
+not get classified. Second, we only ran cross validation on a small subset of the samples.  
+We did not have the time to run it on the full 42,000 samples and thus the k value we used 
+for testing may not have been optimal.
+
 \section{Kernelized Perceptron}
@@
 
-The perceptron algorithm attempts to separate a training data set into two
-classes and attempts to classify future testing samples as belonging to one of
-these two classes. It does this by storing a list of all the mistakes it has
-made previously and using this knowledge to adjust future predictions in the
-hopes of improving classification accuracy.
+The perceptron algorithm is a binary classifier that minimizes the hinge loss objective
+function.  Typically, this would require storing weights to use in making a prediction 
+during testing. However, the kerneled version of the perceptron instead stores the
+mistakes made during training to use during testing. All the mistakes made during
+training are then passed to a kernel with the testing data and test time in order to
+make a prediction of the testing label.
 
@@
 
-One problem that needs to be addressed is that a standard perceptron is a binary
+One problem that needed to be addressed is that a standard perceptron is a binary
 classifier meaning it selects between one of two classes. We resolved this issue
@@
 the polynomial kernel and $\sigma$ in the case of the exponential kernel. We
-used cross validation to find out which kernel gave us the best result as well
-as which value of $d$ and $\sigma$ was optimal.
-
-\begin{table}[ht]
-\textbf{\caption{Perceptron Cross Validation Success Rates}} % title
-\centering
-\begin{tabular}{c | c c }
-Kernel & $d, \sigma = 5$ & $d, \sigma = 10$ \\
-\hline
-Polynomial & 95.67\% & 94.52\% \\
-Exponential & 93.84\% & 96.3\% \\ [1ex]
-\end{tabular}
-\label{table:perceptron}
-\end{table}
+used 10-fold cross validation to find out which kernel gave us the best result as well
+as which value of $d$ and $\sigma$ was optimal. Below are the results of the cross
+validation:
 
 \begin{center}
-\includegraphics[width=406px]{Perceptron_CrossValidation.png}
+\includegraphics[scale=.7]{cv.png}
 \end{center}
+
+There are many different possible selections of kernels to use and our implementation
+allows for the user to pass a command line argument to select the kernel and dimension.
+We chose these two because we knew the performance was adequate based on our
+previous homework. Due to time constraints we were not able to cross validate and expire
+any other kernels.  We briefly explore the Gaussian but the mistake files grew to over 
+1Gb in size and thus the training slowed considerably.
 
@@
 
-Our cross validation indicates that the exponential kernel with a $\sigma$ of 10
+Our cross validation indicates that the exponential kernel with a $\sigma$ of 15
 leads to the highest success rate of 96.3\%. An interesting observation is that
@@
 
-Finally, the runtime of the perceptron algorithm is a little over an hour. It
-takes approximately 45 minutes to train and about 20 minutes to test the
-data. The faster runtime allowed us to cross validate on the full data set so
-our cross validation results are more accurate than those we obtained for KNN.
+Finally, the runtime of the perceptron algorithm is a little over an hour. We timed
+a particular train and test run with the exponential kernel and found that it trained
+in 21 minutes and tested in 48 minutes.  This was similar to what we observers
+with the polynomial kernel as well. The faster runtime allowed us to cross validate 
+on the full data set so our cross validation results are more accurate than those 
+we obtained for KNN.
 
@@
   Nearest Neighbors, $k = 4$   & 96.857\% & $\sim$8 hours\\
-  Perceptron   & 96.300\%  & $\sim$1 hour\\
+  Perceptron   & 96.129\%  & $\sim$1 hour\\
 \end{tabular}
@@
 
-Table 4 summarizes the accuracy and runtimes of weighted KNN with $K = 4$ and
-perceptron with the polynomial kernel and $d = 10$. KNN is clearly more
+Table 4 summarizes the testing accuracy and runtimes of weighted KNN with $K = 4$ and
+perceptron with the exponential kernel and $\sigma = 15$. KNN is clearly more
 accurate, however because perceptron is nearly 8x faster, there could be many
@@
 One reason why KNN may have been so successful in this trial is due to the large
-training data set and small number of classes to classify. With only 10 posible
-classifications, there are likely around 4,2000 training samples of each
-number. The odds of at least a few of those being similar to the test image
-intuitively seem high.
+training data set and small number of classes to classify. With only 10 possible
+classifications, and 42,000 training samples, the odds of at least a few of those
+being very similar to the test image intuitively seem high. As briefly mentioned earlier
+there will be a small minority of strangely written numbers that will not have a good
+match to any of the training samples and thus will not be classified properly.
 
@@
 regularization. This could explain why our polynomial kernel did worse with a
-higher dimension kernel function.
+higher dimension kernel function. We would have liked to explore a kerneled
+SVM to resolve this issue but we ran into problems with derivation and implementation.
+Due to time constraints we stuck with the perceptron and explored its implementation
+further through cross validation and kernel implemenations.
 

### 2013-06-12 Digit Recognition/Kernels.py
Added Cross Validation

@@ -29,3 +29,10 @@ def exponentialKernel(sigma):
         denominator = 2 * sigma ** 2
         return math.exp(-1 * (numerator / denominator))
     return applyKernel
+
+# def exponentialKernel(sigma):
+#     def applyKernel(x_m, x_t):
+#         numerator = np.dot((x_m - x_t), (x_m - x_t))
+#         denominator = 2 * sigma ** 2
+#         return math.exp(-1 * (numerator / denominator))
+#     return applyKernel

### 2013-06-12 Digit Recognition/Perceptron.py

@@ -1,5 +1,4 @@
 __author__ = 'bdwalker'
-
 import numpy as np
 import Kernels
 
@@ -17,7 +16,7 @@ class Classifier:
         #print "Length of x: ", len(x)
         #print "Length of mistakes_x: ", len(mistake_x[0])
 
-        c =  1 if (np.sum(mistake_y * np.apply_along_axis(kernel, 1, mistake_x, x))) > 0 else -1
+        c = 1 if (np.sum(mistake_y * np.apply_along_axis(kernel, 1, mistake_x, x))) > 0 else -1
         return c
 
     # Method to train the model.  It accepts as arguments and ndarray of training data,
@@ -83,4 +82,5 @@ class Classifier:
 
         mistakes = self._trainSVM(x_data[:, 1:], y, report, kernel)
         # Returns length 785
+        print mistakes
         return mistakes

### 2013-06-12 Digit Recognition/PerceptronCrossValidation.py

@@ -1,0 +1,73 @@
+import numpy as np
+from time import time
+import multiprocessing as mp
+import Perceptron
+import Kernels
+import pickle
+import Train
+import Test
+
+
+def crossValidate(train_data):
+    _trainModel(train_data)
+
+def _trainModel(train_data):
+    dimensions = [1, 3, 7, 8, 13, 15, 20]
+    trainer = Train.Trainer()
+    test_size = int(train_data.shape[0] / 10)
+    folds = 10
+    pool = mp.Pool(processes=4)
+    results = list()
+
+    output = open("./results/cv_results_perceptron.txt", "w+")
+    for dim in dimensions:
+        args = (test_size, trainer, folds, train_data, dim)
+        results.append(pool.apply_async(_crossValidate, args))
+
+    pool.close()
+    pool.join()
+
+    for result in results:
+        error, test_time, dim = result.get()
+        output.write("Testing Time: %f" % test_time + " minutes \n")
+        output.write("Polynomial Kernel (dim=" + str(dim) + "):" + str(error / folds) + "\n\n")
+
+    output.close()
+
+def _crossValidate(test_size, trainer, folds, train_data, dim):
+    start = 0
+    error = 0
+    kernel = Kernels.polyKernel(dim)
+    start_time = time()
+    for i in range(folds):
+        split_test = train_data[range(start, start + test_size), 1:]
+        split_label = train_data[range(start, start + test_size), 0]
+        split_train = np.delete(train_data, range(start, start + test_size), axis=0)
+        print "Test Size ", split_test.shape
+        trainer.processData(split_train)
+        model_file = "./Model/perceptron_model_poly_" + str(dim)
+        trainer.trainModel(kernel, True, False, model_file)
+        error += _testModel(split_test, split_label, model_file, kernel)
+        start += test_size
+    end_time = time()
+    return error, (end_time - start_time) / 60.0, dim
+
+def _error(classification, labels):
+    n = labels.shape[0]
+    count = 0
+    for i in range(len(classification)):
+        if int(classification[i]) != int(labels[i]):
+            count += 1
+
+    return float(count)/n
+
+
+def _testModel(test_data, label, model_file, kernel):
+    percep = Perceptron.Classifier()
+    tester = Test.TestModel()
+    inputf = open(model_file, "r")
+    inFile = pickle.load(inputf)
+    model = percep.unpackModel(inFile)
+    classifications = tester.predictWithModel(model, test_data, kernel, True)
+    return _error(classifications, label)
+

### 2013-06-12 Digit Recognition/PerceptronCrossValidation2.py

@@ -1,0 +1,73 @@
+import numpy as np
+from time import time
+import multiprocessing as mp
+import Perceptron
+import Kernels
+import pickle
+import Train
+import Test
+
+
+def crossValidate(train_data):
+    _trainModel(train_data)
+
+def _trainModel(train_data):
+    dimensions = [3, 5, 8, 10, 12, 15, 20]
+    trainer = Train.Trainer()
+    test_size = int(train_data.shape[0] / 10)
+    folds = 10
+    pool = mp.Pool(processes=7)
+    results = list()
+
+    output = open("./results/cv_results_perceptron_expo.txt", "a+")
+    for dim in dimensions:
+        args = (test_size, trainer, folds, train_data, dim)
+        results.append(pool.apply_async(_crossValidate, args))
+
+    pool.close()
+    pool.join()
+
+    for result in results:
+        error, test_time, dim = result.get()
+        output.write("Testing Time: %f" % test_time + " minutes \n")
+        output.write("Exponential Kernel (dim=" + str(dim) + "):" + str(error / folds) + "\n\n")
+
+    output.close()
+
+def _crossValidate(test_size, trainer, folds, train_data, dim):
+    start = 0
+    error = 0
+    kernel = Kernels.exponentialKernel(dim)
+    start_time = time()
+    for i in range(folds):
+        split_test = train_data[range(start, start + test_size), 1:]
+        split_label = train_data[range(start, start + test_size), 0]
+        split_train = np.delete(train_data, range(start, start + test_size), axis=0)
+        print "Test Size ", split_test.shape
+        trainer.processData(split_train)
+        model_file = "./Model/perceptron_model_expo_" + str(dim)
+        trainer.trainModel(kernel, True, False, model_file)
+        error += _testModel(split_test, split_label, model_file, kernel)
+        start += test_size
+    end_time = time()
+    return error, (end_time - start_time) / 60.0, dim
+
+def _error(classification, labels):
+    n = labels.shape[0]
+    count = 0
+    for i in range(len(classification)):
+        if int(classification[i]) != int(labels[i]):
+            count += 1
+
+    return float(count)/n
+
+
+def _testModel(test_data, label, model_file, kernel):
+    percep = Perceptron.Classifier()
+    tester = Test.TestModel()
+    inputf = open(model_file, "r")
+    inFile = pickle.load(inputf)
+    model = percep.unpackModel(inFile)
+    classifications = tester.predictWithModel(model, test_data, kernel, True)
+    return _error(classifications, label)
+

### 2013-06-12 Digit Recognition/Test.py

@@ -7,11 +7,6 @@ import numpy as np
 
 class TestModel:
 
-    def _tallyVotes(self, predictions):
-        lables = [9] * predictions.shape[0]
-        for i in range(predictions.shape[0]):
-            lables[which.predictions[i,:]]
-
     # Passed an unpacked model_dict
     def predictWithModel(self, model_dict, test_data, kernel, report=False):
         perceptron = Perceptron.Classifier()

### 2013-06-12 Digit Recognition/Train.py

@@ -11,7 +11,6 @@ import Perceptron
 
 class Trainer:
     processed_data = None
-    classifiers = None
 
     def _loadProcessedData(self, input_file):
         inputf = open(input_file, "r")
@@ -64,8 +63,6 @@ class Trainer:
                 if key2 > key1:
                     class2 = train_data[key2]
                     class2[:, 0] = -1
-                    print class1.shape
-                    print class2.shape
                     combined = np.vstack((class1, class2))
 
                     # rows need to be shuffled, otherwise not i.i.d
@@ -84,17 +81,22 @@ class Trainer:
         threads = dict()
         cores = multiprocessing.cpu_count()
         pool = mp.ThreadPool(processes=cores)
-        
-        trained_model = dict()
+
+        #trained_model = dict()
         for couple in self.processed_data.keys():
             arg = (self.processed_data[couple], kernel, report, plot)
             result = pool.apply_async(self._trainSingleClass, arg)
             threads[couple] = result
+            #trained_model[couple] = self._trainSingleClass(self.processed_data[couple], kernel, report, plot)
+
+        pool.close()
+        pool.join()
 
         trained_model = dict()
         for r in threads.keys():
-            trained_model[r] = threads[r].get()
-            #Stored length 785
+            thread = threads[r]
+            trained_model[r] = thread.get()
+        #     #Stored length 785
 
         perceptron = Perceptron.Classifier()
         model = perceptron.packModel(trained_model)

### 2013-06-12 Digit Recognition/Main.py
changed main for stat project

@@ -12,20 +12,20 @@ import pickle
 def loadTrainData(input_file):
     return np.loadtxt(input_file, delimiter=",", skiprows=1)
 
-def loadTestData(test_file, true_label=None):
-    test_data = np.loadtxt(test_file, delimiter=",", skiprows=1)
+def loadTestData(test_file, true_label="./uci_data/test/Y_test.txt"):
+    test_data = np.loadtxt(test_file)
     test_y = None
     if true_label is not None:
         print true_label
-        test_y = np.loadtxt(true_label, delimiter=",", skiprows=1)
+        test_y = np.loadtxt(true_label)
     return test_data, test_y
 
-def runSVM():
-    trainer = Train.Trainer()
-    trainer.processData(train_data, None, "./trained_model.csv")
-    kernel = Kernels.exactPolyKernel(5)
-    trainer.trainModel(kernel, True, False, "./trained_model.csv")
-    pass
+def loadUCITrainData(input_x, input_y):
+    train_x = np.loadtxt(input_x)
+    train_y = np.atleast_2d(np.loadtxt(input_y)).T
+    train_data = np.hstack((train_y, train_x))
+    return train_data
+
 
 def runKNNCV(input_file, output, folds):
     train_data = loadTrainData(input_file)
@@ -59,9 +59,9 @@ def runKNN(train_file, test_file, output_file, k, weighted):
 
     print results
 
-def trainPerceptron(train_file, kernel, output_file, dim):
+def trainPerceptron(train_file, kernel, output_file, dim, train_y = None):
     trainer = Train.Trainer()
-    trainer.processData(loadTrainData(train_file))
+    trainer.processData(loadUCITrainData(train_file, "./uci_data/train/Y_train.txt"))
     realKernel = None
     if kernel == "poly":
         realKernel = Kernels.exactPolyKernel(dim)
@@ -87,6 +87,19 @@ def runPerceptron(test_file, kernel, output_file, dim, model_file):
     classifications = tester.predictWithModel(model, testData, realKernel, True)
     output = open(output_file, "w+")
     output.writelines(classifications)
+
+    print calculateError(classifications)
+
+
+def calculateError(classifications):
+    label = np.loadtxt("./uci_data/test/Y_test.txt")
+    error = 0.0
+    for i in range(len(classifications)):
+        if int(classifications[i]) != (label[i]):
+            error += 1.0
+    return error / len(classifications)
+
+
 
 def printUsage():
 

### 2013-06-12 Digit Recognition/Train.py
modified train for stat project

@@ -52,6 +52,7 @@ class Trainer:
             else:
                 train_data[label] = row
 
+        print len(train_data)
         return train_data
 
     def _combineClasses(self, train_data):
@@ -68,7 +69,6 @@ class Trainer:
                     # rows need to be shuffled, otherwise not i.i.d
                     np.random.shuffle(combined)
                     pairings[(key1, key2)] = combined
-
         return pairings
 
     def trainModel(self, kernel=Kernels.defaultKernel, report=False, plot=False, output=None):

### 2013-06-13 Digit Recognition/Main.py
modifications for stat project

@@ -7,7 +7,7 @@ import KNNCrossValidation as cv
 import numpy as np
 import Perceptron
 import pickle
-
+import PerceptronCrossValidation
 
 def loadTrainData(input_file):
     return np.loadtxt(input_file, delimiter=",", skiprows=1)
@@ -99,6 +99,8 @@ def calculateError(classifications):
             error += 1.0
     return error / len(classifications)
 
+def perceptronCV(data_file):
+    PerceptronCrossValidation.crossValidate(loadUCITrainData(data_file, "./uci_data/train/Y_train.txt"))
 
 
 def printUsage():
@@ -172,6 +174,8 @@ def main():
                 dim = int(sys.argv[6])
                 model_file = sys.argv[7]
                 runPerceptron(train_file, kernel, output_file, dim, model_file)
+            elif func == "cv":
+                perceptronCV(train_file)
             else:
                 printUsage()
         else:

### 2013-06-13 Digit Recognition/PerceptronCrossValidation.py

@@ -12,14 +12,14 @@ def crossValidate(train_data):
     _trainModel(train_data)
 
 def _trainModel(train_data):
-    dimensions = [1, 3, 7, 8, 13, 15, 20]
+    dimensions = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 12, 15, 17, 20, 25]
     trainer = Train.Trainer()
     test_size = int(train_data.shape[0] / 10)
     folds = 10
     pool = mp.Pool(processes=4)
     results = list()
 
-    output = open("./results/cv_results_perceptron.txt", "w+")
+    output = open("./results/uci_cv_results_perceptron.txt", "w+")
     for dim in dimensions:
         args = (test_size, trainer, folds, train_data, dim)
         results.append(pool.apply_async(_crossValidate, args))
@@ -45,7 +45,7 @@ def _crossValidate(test_size, trainer, folds, train_data, dim):
         split_train = np.delete(train_data, range(start, start + test_size), axis=0)
         print "Test Size ", split_test.shape
         trainer.processData(split_train)
-        model_file = "./Model/perceptron_model_poly_" + str(dim)
+        model_file = "./Model/uci_perceptron_model_poly_" + str(dim)
         trainer.trainModel(kernel, True, False, model_file)
         error += _testModel(split_test, split_label, model_file, kernel)
         start += test_size
